rgbif: R client for working with GBIF species occurrence data

### 2015-10-27 manuscript.Rmd
updated registry section

`rgbif` is designed following the [GBIF Application Programming Interface][gbifdocs],
or API. The GBIF API has four major components: registry, species names,
occurrence data, and maps. We ignore maps in `rgbif` as it is concerned with
generating maps for web applications. `rgbif` has a suite of functions dealing
with each of registry, species names, and occurrence data - we'll go through
each in turn describing design and example usage.

@@

@@

Search for datasets

@@

Get dataset metrics

```{r}
res <- dataset_metrics(uuid='66dd0960-2d7d-46ee-a491-87b9adcfe7b1')
df <- data.frame(rank = names(res$countByRank),
                 count = unname(unlist(res$countByRank)))
knitr::kable(df)
```

@@

Here, we search for the first give GBIF networks, returning just the
key and title fields.

>>>>

`rgbif` is designed following the [GBIF Application Programming Interface][gbifdocs],
or API. The GBIF API has four major components: registry, species names,
occurrences, and maps. We ignore maps in `rgbif` as it is concerned with
generating maps primarily for web applications. `rgbif` has a suite of functions dealing
with each of registry, species names, and occurrences - we'll go through
each in turn describing design and example usage.

@@

The relationship between these five things is XXXXXX.

@@

Dataset functions include search, dataset metadata retrieval, and dataset metrics.
Searching for datasets is an important part of the discovery process. One can
search for datasets on the GBIF web portal. However, programmatic searching using
this package is much more powerful. Identifying datasets appropriate for a
research question is helpful as you can get metadata for each dataset, and track
down dataset specific problems, if any.

The `dataset_search()` function is one way to search for datasets. Here, we search
for the query term "oregon", which finds any datasets that have terms matching
that term.

@@

Also, check out `datasets()` and `dataset_suggest()` for searching for datasets.

#### Dataset metrics

Dataset metrics are another useful way of digging in to figure out what
datasets you may want to use. One drawback is that these data are only
available for datasets of type _checklist_, but there are quite a lot of
them (`r dataset_search(type = "checklist")$meta$count`).

Here, we seaerch for dataset metrics for a single dataset, with uuid
`ec93a739-1681-4b04-b62f-3a687127a17f`, a checklist of the ants (Hymenoptera:
Formicidae) of the World.

```{r}
res <- dataset_metrics(uuid='ec93a739-1681-4b04-b62f-3a687127a17f')
df <- data.frame(rank = names(res$countByRank),
                 count = unname(unlist(res$countByRank)))
knitr::kable(df)
```

@@

Networks, nodes and installations are at a higher level of organization
above datasets, but can be useful if you want to explore data from given
organizations. Here, we search for the first 10 GBIF networks, returning
just the title field.

### 2015-10-27 manuscript.tex
updated registry section

\texttt{rgbif} is designed following the
\href{http://www.gbif.org/developer/summary}{GBIF Application
Programming Interface}, or API. The GBIF API has four major components:
registry, species names, occurrence data, and maps. We ignore maps in
\texttt{rgbif} as it is concerned with generating maps for web
applications. \texttt{rgbif} has a suite of functions dealing with each
of registry, species names, and occurrence data - we'll go through each
in turn describing design and example usage.

@@

Search for datasets

@@

Get dataset metrics

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{res <-}\StringTok{ }\KeywordTok{dataset_metrics}\NormalTok{(}\DataTypeTok{uuid=}\StringTok{'66dd0960-2d7d-46ee-a491-87b9adcfe7b1'}\NormalTok{)}
\NormalTok{df <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{rank =} \KeywordTok{names}\NormalTok{(res$countByRank),}
                 \DataTypeTok{count =} \KeywordTok{unname}\NormalTok{(}\KeywordTok{unlist}\NormalTok{(res$countByRank)))}
\NormalTok{knitr::}\KeywordTok{kable}\NormalTok{(df)}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[c]{@{}lr@{}}
\toprule
rank & count\tabularnewline
\midrule
\endhead
SPECIES & 52452\tabularnewline
GENUS & 12930\tabularnewline
VARIETY & 4806\tabularnewline
SUBSPECIES & 4440\tabularnewline
SERIES & 1079\tabularnewline
TRIBE & 844\tabularnewline
FAMILY & 509\tabularnewline
SUBTRIBE & 327\tabularnewline
SUBFAMILY & 303\tabularnewline
SUBGENUS & 241\tabularnewline
FORM & 239\tabularnewline
SECTION & 82\tabularnewline
SUBVARIETY & 4\tabularnewline
KINGDOM & 1\tabularnewline
\bottomrule
\end{longtable}

>>>>

\texttt{rgbif} is designed following the
\href{http://www.gbif.org/developer/summary}{GBIF Application
Programming Interface}, or API. The GBIF API has four major components:
registry, species names, occurrences, and maps. We ignore maps in
\texttt{rgbif} as it is concerned with generating maps primarily for web
applications. \texttt{rgbif} has a suite of functions dealing with each
of registry, species names, and occurrences - we'll go through each in
turn describing design and example usage.

@@

Dataset functions include search, dataset metadata retrieval, and
dataset metrics. Searching for datasets is an important part of the
discovery process. One can search for datasets on the GBIF web portal.
However, programmatic searching using this package is much more
powerful. Identifying datasets appropriate for a research question is
helpful as you can get metadata for each dataset, and track down dataset
specific problems, if any.

The \texttt{dataset\_search()} function is one way to search for
datasets. Here, we search for the query term ``oregon'', which finds any
datasets that have terms matching that term.

@@

Also, check out \texttt{datasets()} and \texttt{dataset\_suggest()} for
searching for datasets.

\paragraph{Dataset metrics}\label{dataset-metrics}

Dataset metrics are another useful way of digging in to figure out what
datasets you may want to use. One drawback is that these data are only
available for datasets of type \emph{checklist}, but there are quite a
lot of them (2472).

Here, we seaerch for dataset metrics for a single dataset, with uuid
\texttt{ec93a739-1681-4b04-b62f-3a687127a17f}, a checklist of the ants
(Hymenoptera: Formicidae) of the World.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{res <-}\StringTok{ }\KeywordTok{dataset_metrics}\NormalTok{(}\DataTypeTok{uuid=}\StringTok{'ec93a739-1681-4b04-b62f-3a687127a17f'}\NormalTok{)}
\NormalTok{df <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{rank =} \KeywordTok{names}\NormalTok{(res$countByRank),}
                 \DataTypeTok{count =} \KeywordTok{unname}\NormalTok{(}\KeywordTok{unlist}\NormalTok{(res$countByRank)))}
\NormalTok{knitr::}\KeywordTok{kable}\NormalTok{(df)}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[c]{@{}lr@{}}
\toprule
rank & count\tabularnewline
\midrule
\endhead
SPECIES & 13710\tabularnewline
SUBSPECIES & 3234\tabularnewline
GENUS & 726\tabularnewline
TRIBE & 53\tabularnewline
SUBFAMILY & 20\tabularnewline
FAMILY & 2\tabularnewline
KINGDOM & 1\tabularnewline
PHYLUM & 1\tabularnewline
CLASS & 1\tabularnewline
ORDER & 1\tabularnewline
\bottomrule
\end{longtable}
